---
date: "`r Sys.Date()`"
params:
  input_rds:
    value: x
  output_rds:
    value: x
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: yes
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 6, cache.lazy = FALSE)
```

# Set-Up

## Libraries

```{r setup_libs}
library(Seurat)
library(data.table)
library(ggplot2)
library(knitr)
library(networkD3)
library(patchwork)
library(dplyr)
library(clustree)
library(ggthemes)
library(viridis)
library(yaml)
library(tibble)
library(future)
library(plotly)
library(RColorBrewer)
library(parallel)
```

## Variables

## Update according to your dataset

```{r setup_variables}

# read config file in yaml format
if (!file.exists("config/config.yaml")) {
  stop("config/config.yaml does not exist. Exiting the program")
} else {
  config_file <- read_yaml("config/config.yaml")
}

doc_title <- paste(config_file$title, "- clustering")
author_list <- paste(config_file$authors, collapse = ", ")

# Clustering Resolution

clusRes <- lapply(names(config_file$resolutions), function(x) {config_file$resolutions[[x]]}) # resolution values (floats)
names(clusRes) <- names(config_file$resolutions) # map sample name to resolution

# Number of dimensions to use for individual samples
samplePC <- config_file$samplePC

# Number of dimensions to use for integrated samples
integratedPC <- config_file$integratedPC

# check if resolutions > 2
if (any(clusRes > 2)) {
  stop("Cluster resolution is > 2")
}

ab_list <- config_file$antibody_list

feature_plot_list <- config_file$FeaturePlotList
```

---
title: "`r doc_title`"
author: "`r author_list`"
---

## Inherent Variables

```{r setup_inherent_variables}

# Load Seurat objects
exptsList <- readRDS(params$input_rds)

# Create plot that explains variance by Principle Components (PCs) to find the dimensionality
percentExplained <- function( seuratObject ){
  return(cumsum(seuratObject@reductions$pca@stdev^2)/seuratObject@reductions$pca@misc$total.variance)
}

plotPercentExplained <- function( seuratObject , ndims = nDims) {
  data.use <- percentExplained( seuratObject )
  plot <- ggplot(data = data.frame(dims = 1:ndims, var = data.use[1:ndims])) +
            geom_point(mapping = aes_string(x = 'dims', y = 'var')) + labs( y = "% variance explained", x = "PC") +
            ggtitle(sprintf("Variance Explained by Principle Components in %s", sample)) +
            theme_clean()
  return(plot)
}
```

# Clustree Clustering

```{r clustree, results = 'hide', message = FALSE}

# compute neighbors and perform clustree analyses per sample
plotList <- list()

for (sample in names(exptsList)) {
  
  print(paste("Clustree analysis for sample", sample))
  myso <- exptsList[[sample]]
  
  if (sample == "integrated") {
    prefix = "integrated_snn_res."
    nDims = integratedPC
  } else {
    prefix = "SCT_snn_res."
    nDims = samplePC
  }

  myso <- myso %>%
    FindNeighbors(dims = 1:nDims) %>%
    FindClusters(resolution = seq(0.1, 2, 0.1))

  # Add UMAP embedding
  embeddings <- Embeddings(object = myso, reduction = "umap")

  myso@meta.data$UMAP_1 <- embeddings[,1]
  myso@meta.data$UMAP_2 <- embeddings[,2]

  # Clustree visualizations to determine best cluster resolution
  plotList[[sample]][["clustree"]] <- clustree(myso, prefix = prefix)
  
  plotList[[sample]][["clustreeOverlay"]] <- clustree_overlay(x = myso@meta.data,
                                                              prefix = prefix,
                                                              x_value = "UMAP_1",
                                                              y_value = "UMAP_2")

  myso@meta.data$UMAP_1 = NULL
  myso@meta.data$UMAP_2 = NULL

  exptsList[[sample]] <- myso
}

print('1')
```

# UMAP Parameters per Sample {.tabset}

```{r umap_params, results='asis'}
for (i in names(exptsList)) {
  cat(paste( "##", toupper(i), "\n\n" ))

  cat("### Clustree Hierarchy\n")
  plot(plotList[[i]][["clustree"]])
  cat("\n\n---\n\n")
 
  cat("### Clustree Overlay\n")
  plot(plotList[[i]][["clustreeOverlay"]])
  cat("\n\n---\n\n")
}
```

# UMAP Clustering per Sample

```{r umap}

for (sample in names(exptsList)) {
  
  paste(sprintf("UMAP Clustering the %s object", sample))
  
  myso <- exptsList[[sample]]
  
  # Cluster the cells at the designated resolution in the config file.
  myso <- myso %>% FindClusters(resolution = clusRes[[sample]], verbose = FALSE)
  
  # Plot and save UMAP clusters
  Idents(myso) <- myso$seurat_clusters
  plotList[[sample]][["umap"]] <- DimPlot(object = myso, reduction = 'umap', label = TRUE, label.size = 7, pt.size = 1) + ggtitle(toupper(sample))
  
  # Save clustering
  exptsList[[sample]] <- myso
}

rm(myso)
```

# UMAP Plots {.tabset}
```{r umap_plots, results='asis'}
for (i in names(exptsList)) {
  cat(paste( "##", toupper(i), "\n" ))

  cat("\n")
  plot(plotList[[i]][["umap"]])
  cat("\n\n")
}
```

# Custom Feature Plots {.tabset}
```{r rnaFeaturePlot, results='asis'}

# Fetch list of viable genes per each sample (subscriptable_genes).
# Remove any genes from original list from subscriptable_genes.
# Append to filtered list of genes, then unlist and unique.
feature_plot_list_filtered <- list()
for (sample in names(exptsList)) {
  DefaultAssay(exptsList[[sample]]) <- "SCT"
  subscriptable_genes <- colnames( FetchData(exptsList[[sample]], vars = c("ident", feature_plot_list)) )
  subscriptable_genes <- try({feature_plot_list[feature_plot_list %in% subscriptable_genes]}, silent = TRUE)
  feature_plot_list_filtered <- append(feature_plot_list_filtered, subscriptable_genes)
}
feature_plot_list_filtered <- unique(unlist(feature_plot_list_filtered))

for (sample in names(exptsList)) {
  cat(paste("##", toupper(sample), "{.tabset}", "\n\n"))
  DefaultAssay(exptsList[[sample]]) <- "RNA"

  for (gene in feature_plot_list_filtered) {
    cat(paste("###", gene, "\n\n"))

    cat("\n\n")
    plot(FeaturePlot(exptsList[[sample]], gene, pt.size = 1, label = TRUE, label.size = 8))
    cat("\n\n")
  }
}
print('2')
```

# Cluster Proportions by Sample

```{r clusterProp}
exptsList[["integrated"]]@meta.data %>%
  group_by(seurat_clusters, orig.ident) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = seurat_clusters, y = prop, fill = orig.ident)) +
  geom_col() +
  ggtitle("Cell proportions by cluster") +
  ylab("Proportions") +
  labs(fill = "Treatment") +
  geom_text(aes(label = round(prop * 100, 1)), position = position_stack(vjust = 0.5), color = "white", size = 4)

exptsList[["integrated"]]@meta.data %>% 
    group_by(seurat_clusters,orig.ident) %>% 
    summarise(n = n()) %>%
    ungroup() %>%
    group_by(orig.ident) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(x = orig.ident, y = prop, fill = seurat_clusters)) +
    geom_col() +
    ggtitle("Cluster Proportions by Sample") + 
    geom_text(aes(label = round(prop * 100, 1)), position = position_stack(vjust = 0.5), color = "white", size = 4)
```

# 3D UMAP Colored by Clusters

```{r 3D_umaps}

# define umap output directory
if (!dir.exists("data/umaps")) {
  dir.create("data/umaps")
}

for (sample in names(exptsList)) {

  myso <- exptsList[[sample]]

  cell_meta <- merge(myso@reductions$umap@cell.embeddings, myso@meta.data, by = 0, all = TRUE)
  cell_meta <- cell_meta %>% column_to_rownames("Row.names")
  col_scheme <- colorRampPalette(brewer.pal(8, config_file$rcolorbrewer_palette))(length(unique(cell_meta$seurat_clusters)))

  p <- plot_ly(cell_meta,
          x = ~UMAP_1,
          y = ~UMAP_2,
          z = ~UMAP_3,
          size = 1,
          color = ~seurat_clusters,
          colors = col_scheme,
          # hover text
          text = ~paste("Cluster:", seurat_clusters, "<br>nCount_SCT:", nCount_SCT)) %>% 
    add_markers() %>%
    layout(title = paste(toupper(sample), "UMAP Clusters"),
          xaxis = list(title = "UMAP_1"),
          yaxis = list(title = "UMAP_2"))

  # save and print where they are located.
  out_file <- paste0("data/umaps/", toupper(sample), "_umap.html")
  print(paste("Sample", sample, "UMAP file is located at:", out_file))
  htmlwidgets::saveWidget(p, out_file) # Rmd does not output files.
}

rm(myso)
```

# 3D UMAP Colored by Identity

```{r 3D_umaps2}

if (!dir.exists("data/umaps")) {
  dir.create("data/umaps")
}

myso <- exptsList[["integrated"]]

cell_meta <- merge(myso@reductions$umap@cell.embeddings, myso@meta.data, by = 0, all = TRUE)
cell_meta <- cell_meta %>% column_to_rownames("Row.names")
col_scheme <- colorRampPalette(brewer.pal(8, config_file$rcolorbrewer_palette))(length(unique(cell_meta$seurat_clusters)))

p <- plot_ly(cell_meta,
        x = ~UMAP_1,
        y = ~UMAP_2,
        z = ~UMAP_3,
        size =1,
        color = ~orig.ident,
        colors = col_scheme,
        # hover text
        text = ~paste("Cluster:", seurat_clusters, "<br>nCount_SCT:", nCount_SCT)) %>% 
  add_markers() %>%
  layout(title = "Integrated UMAP by Sample Identity",
        xaxis = list(title = "UMAP_1"),
        yaxis = list(title = "UMAP_2"))

out_file <- "data/umaps/integrated_identity_umap.html"
print(paste("Integrated UMAP colored by sample identity is here:", out_file))
htmlwidgets::saveWidget(p, out_file)

rm(cell_meta, col_scheme, p, myso, out_file)

```

# Save Data
```{r save_data, include=TRUE}
print(paste("Saving preprocessed individual samples and the integrated object in", params$output_rds))
saveRDS(exptsList, file = params$output_rds)
```

# Session Information
```{r session_info}
sessionInfo()
```

